# R markdown

## Headings

Every chapter should have a first-level heading marked by `#`. A second-level heading has two of `#`, a third-level heading three of them and so on until six. Here's how to write a first-level heading followed by a second-level heading:

````markdown
# <first-level heading>

## <second-level heading>

````

## Paragraphs

Paragraphs are separated by two line-breaks:

````{verbatim}
<one paragraph>

<next paragraph>

````

## Italics

Text can be turned to italics using asterisks around it:

````markdown
*<text in italics>*

````

*\<text in italics\>*

## Code

### As a block

Code blocks can be marked by three ticks following by braces containing metadata. The first metadata is the language. i've used so far `r` and `bash`:

````{verbatim}
```{bash eval = FALSE}
ssh-keygen

```

````

i keep an empty line at the end of the code block content because if i want to use a keyboard shortcut for commenting out it only works if the last line is empty.

There are some options that can be used for code blocks or chunks in addition to them mentioned in the general [configuration](#knitr) part:

`eval`
: Shall the code be executed or not or what lines of the code should be executed?

### Inline

The inline code must be between ticks:

````{verbatim}
<text before inline code>`<inline code>`

````

\<text before inline code\>`<inline code>`

## Quoting

For quoting, i use `>` at the beginning of each line of the quote.

````{verbatim}
> <the quoted text>

````

> \<the quoted text\>

## Escaping

If a character must be escaped then a backslash can be used right in front of it:

````markdown
*\**

````

*\**

## References

A target for a common reference:

````markdown
[]{#<label-for-target>}

````

A link for a common target:

````markdown
[<link text>](#<label-for-target>)

````

## Figures

For figures, i created a wrapper function `include_external_graphics()` for `knitr`'s `include_graphics()` because natively, `include_graphics()` doesn't support all the necessary file types: 

````{verbatim}
```{r label = "<label of figure>", fig.cap = "<caption of figure>"}
include_external_graphics("<path-to-image-file>")

```

````

For this code chunk, i use the label and the following options:

fig.cap
: the caption of the figure. This must be present in order to reference it.

Referencing a figure takes place using `\@ref(fig:<label of figure>)`.

As an example, i refer to the figure \@ref(fig:workflow-example).

```{r label = "workflow-example", fig.cap = "The workflow from the written markdown to the ready output using *bookdown*."}
include_external_graphics("rmd/workflow.png")

```
